---
title: "Marketing Analytics"
author: "Sanjid Rahman"
date: "26/11/2021"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library, include=FALSE}
library(data.table); library(lubridate); library(tidyr); library(dplyr)
library(ggplot2); library(gridExtra); library(cowplot)
```


```{r load_data, include=FALSE}
#df <- fread("data/marketing_data.csv")
df<- read.csv("data/marketing_data.csv")
#head(as.data.table(df), 6)
```

```{r data_cleaning, echo=FALSE}
#df[, Dt_Customer := as.Date(parse_date_time(Dt_Customer,"mdy"))]
df$Dt_Customer<-as.Date(parse_date_time(df$Dt_Customer,"mdy"))
#df[, Income := as.numeric(gsub("[$]|[,]", "", Income))]
df$Income<-as.numeric(df$Income%>%gsub("[$,]","",.))
#df <- df[Year_Birth < 1920]
df<-mutate(df, Age=2020-df$Year_Birth)
#paste0("Number of Datapoints removed for Age>100: ",count(df%>%filter(.,Age>100)))
df <-df %>% filter(., Age<100)
#summary(df$Age)
df<-df%>%filter(.,!is.na(Income))
```

## Existing features

`r names(df)`

`r dim(df)[1]` rows and `r dim(df)[2]` columns

## New Features

- Total Amount Spent = MntWines + MntFruits + MntMeatProducts + MntFishProducts + MntSweetProducts + MntGoldProds
- Total Campaigns Accepted = AcceptedCmp1 +  AcceptedCmp2  + AcceptedCmp3 +  AcceptedCmp4 +  AcceptedCmp5
- Total Number Of Purchases = NumDealsPurchases + NumCatalogPurchases + NumStorePurchases + NumWebPurchases
- Number of Dependents = Kidhome +  Teenhome
- Number of Years as Customers = 2020- Dt_Customer

```{r new_features, echo=FALSE}
df<-df%>%mutate(df,Total_amt_spent= MntWines+ MntFruits+ MntMeatProducts+ MntFishProducts+ MntSweetProducts+ MntGoldProds)
#print("Total_amt_spent")
#summary(df$Total_amt_spent)

#Total Campaigns Accepted
df<-df%>%mutate(df,Num_campaigns_Accepted= AcceptedCmp1+ AcceptedCmp2 +AcceptedCmp3+ AcceptedCmp4+ AcceptedCmp5)
#print('Num_campaigns_Accepted')
#summary(df$Num_campaigns_Accepted)

#Total Number of Purchases
df<-df%>%mutate(df,Num_purchases= NumDealsPurchases+ NumCatalogPurchases+ NumStorePurchases+ NumWebPurchases)
#print("Num_purchases")
#summary(df$Num_purchases)

#Number of Dependents
df<-df%>%mutate(df,Num_dependents= Kidhome+ Teenhome)
#print("Num_dependents")
#summary(df$Num_dependents)

#Num_year joined if current year 2020
df<-df%>%mutate(df, Num_years= 2020- year(Dt_Customer))
#print("Num_year_joined")
#summary(df$Num_years)
```


```{r message= FALSE, warning= FALSE,error=FALSE}
df_regress<-df #%>%select(-ID,-Year_Birth,-Dt_Customer)
```
## Overview of the Y-Variable : Num Store Purchase
```{r message= FALSE, warning= FALSE,error=FALSE}
summary(df$NumStorePurchases)
p1<-df_regress%>%ggplot(aes(NumStorePurchases))+
  geom_histogram(fill="#666699")+
  scale_x_continuous(n.breaks=13)+
  ggtitle("Frequency of Number of Store Purchases")+
  ylab("Frequency")  

p2<-df_regress%>%ggplot(aes(y=NumStorePurchases))+ 
  geom_boxplot(fill="#666699")+ 
  theme(axis.ticks.x=element_blank(),
        axis.text.x = element_blank())

grid.arrange(p1,p2,ncol=1)
```

## Visualize Correlation
```{r correlation, message= FALSE, warning= FALSE,error=FALSE}
library(caret)
df1<-select_if(df_regress,is.numeric)
correlation_matrix<-cor(df1)

#Visualise correlation in heatmap
col<- colorRampPalette(c("blue", "white", "red"))(20)
heatmap(correlation_matrix, col=col, symm=TRUE)   
```


- Highly positively correlated variables are in Red and Highly negatively correlated variables are in Blue.
- We could see that income are positively correlated to Number of purchases and the amount of purchases.
- The different kind of purchases such as Meat purchases/ Sweet purchases/ Fish purchases/ Fruit purchases tend to be positively correlated to one another.
- There are negative correlations between having kids/ dependents at home and the amount or number of purchases.


#### Drop Highly correlated variables  
```{r message= FALSE, warning= FALSE,error=FALSE}
drop=findCorrelation(correlation_matrix,cutoff=0.8)
drop= c(drop)
corrdf<-data.frame(correlation_matrix)
col_drop<-corrdf%>%dplyr::select(drop)%>%colnames()
df_regress<-df_regress%>%select(-col_drop)
paste0("Columns dropped: ", col_drop)  
```
  
  
#### Numeric labeling: Convert to numerical representation
```{r message= FALSE, warning= FALSE,error=FALSE}
#retrieve all non-numeric columns and represent these factors as integers.
library(tidyverse)
cat_var<-df_regress%>%select_if(negate(is.numeric))%>%colnames()
paste0("Categorical Variables : ", cat_var)
## [1] "Education"      "Marital_Status" "Country"

df_regress$Education1<-as.integer(factor(df_regress$Education))
df_regress%>%distinct(Education,Education1)

df_regress$Marital_Status1<-as.integer(factor(df_regress$Marital_Status))
df_regress%>%distinct(Marital_Status,Marital_Status1)

df_regress$Country1<-as.integer(factor(df_regress$Country))
df_regress%>%distinct(Country,Country1)

df_regress1<-df_regress%>%select(-Education,-Marital_Status,-Country,-Num_campaigns_Accepted,-Num_dependents)  
```
  
  
#### Split data into training (70%) & test datasets (30%)
```{r message= FALSE, warning= FALSE,error=FALSE,echo=FALSE}
library(caTools)
set.seed(123)
df_regress1$spl<-sample.split(df_regress1$NumStorePurchases,SplitRatio = 0.7)
train_df<-df_regress1%>%filter(spl==TRUE)%>%select(-spl)
test_df<-df_regress1%>%filter(spl==FALSE)%>%select(-spl)
paste0("Number of rows in train dataset: ",nrow(train_df))
paste0("Number of rows in test dataset: ",nrow(test_df))
```
  
  
#### Perform Linear Regression & Interpret Results
  
#### Regression Results
```{r message= FALSE, warning= FALSE,error=FALSE,max.height='300px', echo=FALSE}
m1<-lm(NumStorePurchases~.,data=train_df)
summary(m1)
```
  
  
#### Plotting of all X-Variables' P-value
```{r message= FALSE, warning= FALSE,error=FALSE,echo=FALSE}
#Retrieve all the p-values
pvaluedf<-data.frame(summary(m1)$coefficients[,c('Pr(>|t|)','Estimate')])
colnames(pvaluedf)<-c('pvalue','coefficient')
pvaluedf$variables<-rownames(pvaluedf)

#Plot the variables and their significance 
pvaluedf%>%ggplot(aes(x=reorder(variables,pvalue),y=pvalue))+
  geom_col(fill="#666699")+
  theme(axis.text.x=element_text(angle=90,hjust=1))+
  geom_hline(yintercept=0.05,color="red")+
  geom_text(aes(x=2,y=0.10),label="5%",color="red")+
  scale_x_discrete(name="X-Variables")+
  ggtitle("X-Variables and their P-Values")


```
  
#### The table below shows the top 5 significant X-variables for NumStorePurchase
```{r message= FALSE, warning= FALSE,error=FALSE, echo=FALSE}
#Top 5 significant variables based on p-values (significance level at 5%) 
pvaluedf%>%arrange(pvalue)%>%select(pvalue,coefficient)%>%slice(1:6)

#Calculate RMSE Value
predictions<-predict(m1,test_df)
paste0("RMSE: ",RMSE(predictions,test_df$NumStorePurchases))

#R-square on training dataset
paste0("R-square value on train dataset is: " ,round(summary(m1)$r.squared,digits=5))

#Error Sum of Squares
SSE<- sum((predictions-test_df$NumStorePurchases)^2)
#Sum of Square Total
SST<- sum((mean(train_df$NumStorePurchases)-test_df$NumStorePurchases)^2)
#R-Square 
r_square<-1-SSE/SST
paste0("R-square value on test dataset is: ", round(r_square,5))

```
  
#### Plot the 5 significant X-variables with the Y-variable
```{r message= FALSE, warning= FALSE,error=FALSE,echo=FALSE}
library(gtable)
p1<-df_regress1%>%
  select(NumStorePurchases, NumWebPurchases, NumWebVisitsMonth, NumDealsPurchases)%>%
  gather(.,var,value,'NumWebPurchases':'NumDealsPurchases')%>%
  group_by(NumStorePurchases,var)%>%
  summarise_all(funs(mean))%>%
  ggplot(aes(x=NumStorePurchases,y=value,fill=var))+
  geom_col(position="dodge2")+ 
  scale_x_continuous(n.breaks = 11)+
  theme(legend.position="bottom",
        plot.title = element_text(size = 10,face = "bold"),
        text=element_text(size=9))+
  scale_fill_brewer(palette = "Dark2")+
  ggtitle("Mean number of X-variables and NumStorePurchases")

p2<-df_regress1%>%
  select(NumStorePurchases,MntWines)%>%
  group_by(NumStorePurchases)%>%
  summarise(MntWines=mean(MntWines))%>%
  ggplot(aes(x=NumStorePurchases,y=MntWines))+
  geom_col(fill="#666699",width=0.4)+
  theme(legend.position="none")+
  scale_x_continuous(n.breaks = 11)+
  scale_y_continuous(label=scales::dollar)+
  ggtitle("Mean amount of Wine Purchases and NumStorePurchases")+
  theme(plot.title = element_text(size = 10,face = "bold"),
        text=element_text(size=9))

p3<-df_regress1%>%
  ggplot(aes(x=factor(Kidhome),y=NumStorePurchases))+
  geom_boxplot()+
  scale_x_discrete(limits=c("0", "1","2"))+ 
  xlab("Number of Kids")+
  ggtitle("Number of kids and \n Mean NumStorePurchases")+
  theme(plot.title = element_text(size = 10,face = "bold"),
        text=element_text(size=9))

grid.arrange(p2,p3,p1, ncol=2,widths=c(2,1),heights=c(1.5,2))
```
  
#### Observation:
- Amount spent on Wines increases as the Number of Store Purchases increases
- Number of Web Purchases increases as the Number of Store Purchases increases
- Number of Web Visits decreases as the Number of Store Purchases increases
- Mean Number of Store Purchases decreases when there are kids
- These observations suppors the regression model result which indicates positive coefficients for MntWines, NumWebPurchases and negative coefficients for KidHome and NumWebVisitsMonth
  

---